Simulations of Ballot Polling Risk-Limiting Audits

### 2023-02-13 providence/sections/intro.tex
added margin and bravo references to beta

%Intro
It is well-known that electronic voting systems are vulnerable to software errors and manipulation which may be undetected. Errors and/or manipulation may not always change an election outcome, but we would want to know when they do. {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome. {\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; the evidence is examined so that incorrect outcomes may be detected with high probability. One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots.

@@

\subsection{Background on RLAs}
All RLAs sample one or more ballots at random; we will refer to each such set of ballots as a {\em round}. The ballots are manually examined, and a stopping condition computed, which determines whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn (more information is needed before a determination). In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. However, a manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary, hence (c) is generally not incorporated. Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds. They would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.

@@

Ballot polling audits proceed as follows. 
\begin{enumerate}
\item A first round \cite{usenix_minerva} size---the number of ballots sampled before first checking the stopping condition---is chosen. 
\item Ballots on the ballot manifest are sampled uniformly at random using a pseudorandom number generator typically seeded by a natural source of randomness like rolling dice.
\item The physical ballots are found and manually interpreted, recording the manual interpretations. 
\item Based on the manual interpretations, the stopping condition is computed. 
\item If more ballots are to be drawn, the next round size is chosen. Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required. For RLAs required by statute or law, certification deadlines would play a large role, as the audit would need to be completed before the deadline. 
\end{enumerate}

A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots; typically hundreds or thousands or tens of thousands of ballots in statewide elections. A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn. The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).

Election officials typically draw ballots in large round sizes:  see for example \cite{va-2022,RI-report}. Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping). Further, at both audits we attended, election officials chose stopping probabilities of $0.9$ and we are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds). \BRAVO hence cannot be used as a B2 audit in these scenarios.

>>>>

%Intro
It is well-known that electronic voting systems are vulnerable to software errors and manipulation which may be undetected. Errors and/or manipulation may not always change an election outcome, but we would want to know when they do. {\em Software independent} voting systems \cite{SI-Wack,rivest2008notion} are ones where an undetected change in the software cannot lead to an undetectable change in the election outcome. {\em Evidence-based elections} \cite{evidence-based} use software independent systems to produce trustworthy evidence of outcome correctness; the evidence is examined so that incorrect outcomes may be detected with high probability. One approach to evidence-based elections is to use voter-verified paper ballots, store them securely, and perform public audits---a compliance audit to determine whether the ballots were stored securely; and a rigorous tabulation audit, known as a risk-limiting audit (RLA) \cite{RLA}, to determine whether the outcome is correctly computed from the stored ballots. To have any value, the audits should be performed before the election is certified.

@@

\subsection{Background on RLAs}
All RLAs sample one or more ballots at random; we will refer to each such set of ballots as a {\em round}. The ballots are manually examined, and a stopping condition computed, which determines whether (a) the audit ends in success (the election outcome is declared correct) or (b) another round should be drawn (more information is needed before a determination). In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count, which is generally not incorporated \cite{bravo}. A manual hand count presents significant logistical challenges, and there is always a chance that it will have been unnecessary. Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds. They would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence.

@@

Ballot polling audits proceed as follows. 
\begin{enumerate}
\item A first round \cite{usenix_minerva} size---the number of ballots sampled before first checking the stopping condition---is chosen. 
\item Ballots on the ballot manifest are sampled uniformly at random using a pseudorandom number generator typically seeded by a natural source of randomness like rolling dice.
\item The physical ballots are found and manually interpreted, recording the manual interpretations. 
\item Based on the manual interpretations, the stopping condition is computed. 
\item If more ballots are to be drawn, the next round size is chosen, and we go back to step 2. 
\end{enumerate}

Round sizes, including the first one, may be computed based on a desired probability of audit completion at the end of the round, and may take into consideration loose estimates of the resources required. For RLAs required by statute or law, certification deadlines would play a large role, as the audit would need to be completed before the deadline.

A {\em round-by-round (R2)} audit is the general audit, where the decision of whether to draw more ballots or not is taken after drawing a round of ballots. A {\em ballot-by-ballot (B2)} audit is the special case of round size one---when the decision is made after each ballot is drawn. The popular \BRAVO audit requires the smallest expected number of ballots when the announced tally of the election is correct, and stopping decisions are taken a ballot at a time (that is, when it is used as a B2 audit).

Election officials typically draw ballots in large round sizes, see for example \cite{va-2022,RI-report}. Note also that, in addition to allowing users to directly enter a round size or choose the expected number of ballots drawn by \BRAVO, Arlo provides choices of stopping probabilities of $0.9$, $0.8$ and $0.7$ (\cite{usenix_minerva} provides estimates of these round sizes for each state in the 2020 US Presidential election, for a $0.9$ probability of stopping; some numbers are in tens and hundreds of thousands of ballots, and thousands of ballots are quite common). Further, at both audits we attended, election officials chose stopping probabilities of $0.9$ and we are not aware of any ballot polling RLA performed on ballots cast in a governmental election that drew ballots one at a time (though the stopping condition can be computed one ballot at a time, the ballots are drawn in rounds). \BRAVO hence cannot be used as a B2 audit in these scenarios.

### 2023-02-13 providence/sections/model.tex
added margin and bravo references to beta

An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}. As described in the introduction, we ignore a third option of moving to a hand count, leaving that decision to election officials. We provide more context on this in section \ref{sec:related}, on related work.

@@

>>>>

An audit $\mathcal{A}$ is a function that takes as input the sample of ballots and outputs either (1) \emph{Correct: stop the audit} or (2) \emph{Undetermined: sample more ballots}. As described in the introduction, we ignore a third option of moving to a hand count, leaving that decision to election officials. This is standard in the literature, see, for example, \cite{bravo}. We provide more detail on the use of this option in section \ref{sec:related}, on related work.

@@

Notice that the ratios for \BRAVO and \Minerva are for two candidates. They may both be extended to multi-candidate plurality elections through the computation of $\ell -1$ audits for $\ell$ candidates by pairing the announced winner with all other candidates.

The {\em margin}, given a contest and two candidates, is the difference between the votes received by each as a fraction of the total number of votes.

### 2023-02-13 providence/sections/related.tex
added margin and bravo references to beta

In the existing literature, ballot polling audits do not include this possibility (i.e. they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.

@@

For completeness, we describe here other types of RLAs. 
\begin{description}
\item In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot. Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot. A typical approach is to use a ballot serial number on both paper ballot and CVR. When voters vote in precincts, however, serial numbers on ballots can enable the correlation of ballots with voters, and ballots are typically not numbered. Additionally, some voting systems do not record a CVR. One may perform a transitive audit by rescanning unnumbered voted ballots with special scanners which produce CVRs and also print numbers on the ballots as they are scanned. This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible. 
\item A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch. Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch. This approach typically requires the sampling of a very large number of ballots. However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied. (Note that, for an RLA, the number of batches tallied is not fixed; the risk limit is. A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
\end{description}

>>>>

An example of the above use is \cite{RLA}. In the more recent literature, for example \cite{bravo}, ballot polling audits do not include this possibility (i.e. they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count.

@@

For completeness, we describe other types of RLAs below. 
\begin{description}
\item In a ballot comparison RLA \cite{RLA}, the manual interpretation of each sampled ballot is compared to the corresponding Cast Vote Record (CVR), which is the machine interpretation of the ballot. Ballot comparison RLAs require the fewest ballots of all known RLA approaches, but also require a means of identifying the CVR corresponding to a particular ballot. A typical approach is to use a ballot serial number on both paper ballot and CVR. When voters vote in precincts, however, serial numbers on ballots can enable the correlation of ballots with voters, and ballots are typically not numbered. Additionally, some voting systems do not record a CVR. One may perform a transitive audit by rescanning unnumbered voted ballots with special scanners which produce CVRs and also print numbers on the ballots as they are scanned. This requires an investment in a sufficiently large number of such printers and in the human effort of rescanning ballots, and is not always feasible. 
\item A batch comparison RLA \cite{RI-report} samples batches of ballots (typically, a batch is a storage box of ballots) and compares the manual tally of each sampled batch with the announced tally of that batch. Thus, while this type of audit does not need CVRs, it does need both a ballot manifest and a public declaration of the tally of each batch. This approach typically requires the sampling of a very large number of ballots. However, the process is most similar to that which election officials already use when they perform fixed-percentage post-election audits, where, for example, 2\% of the batches are manually tallied. (Note that, for an RLA, the number of batches tallied is not fixed; the risk limit is. A smaller number of batches might be sufficient, or a larger number necessary, for an audit with the required risk limit.)
\end{description}
